---
title: "In-Class Lab 16"
author: "ECON 4223 (Prof. Tyler Ransom, U of Oklahoma)"
date: "April 11, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', fig.keep = 'none')
```

The purpose of this in-class lab is to use R to practice with panel data estimation methods. The lab should be completed in your group. To get credit, upload your .R script to the appropriate place on Canvas.

## For starters
Open up a new R script (named `ICL16_XYZ.R`, where `XYZ` are your initials) and add the usual "preamble" to the top:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Add names of group members HERE
library(tidyverse)
library(wooldridge)
library(broom)
library(magrittr)
library(stargazer)
library(clubSandwich)
library(plm)   # You may need to install this package
```

### Load the data
Our data set will be a panel of wages for 545 men. Load the data from the `wooldridge` package, format `year` to be a factor, and rename the variable `nr` to something more descriptive (`id`):
```{r}
df <- as_tibble(wagepan)
df %<>% mutate(year=as.factor(year))
df %<>% rename(id = nr)
```

### Summary statistics for panel data
It is important to know what your panel looks like. How many units? How many time periods? The easiest way to do this is the `pdim()` function in the `plm` package.
```{r}
pdim(df)
```

It is also helpful to "convert" the data to a cross-section of within-unit averages. Let's do this for some of the key variables of our analysis.
```{r}
df.within <- df %>% select(id,year,educ,married,union,rur) %>% 
             group_by(id) %>% 
             summarize(
                 mean.edu = mean(educ),
                 var.edu  = var(educ),
                mean.marr = mean(married),
                 var.marr = var(married),
               mean.union = mean(union),
                var.union = var(union),
               mean.rural = mean(rur),
                var.rural = var(rur)
             )
df.within %>% as.data.frame %>% stargazer(type="text")
```

1. Is there any within-person variance in the `educ` variable? What about `married`, `union`, and `rural`?

2. What does it mean for the married, union, or rural variables to have a positive within-person variance?

3. Why is it important to know if a variable has positive within-person variance?

## Pooled OLS, Random Effects, and Fixed Effects Models
Now estimate the following model using various options of the `plm()` function:

\begin{align*}
\log(wage_{it}) & = \beta_0 + \beta_1 educ_{i} + \beta_2 black_{i} + \beta_3 hisp_{i} + \beta_4 exper_{it} + \beta_5 exper_{it}^2 + \beta_6 married_{it} + \\ 
&\phantom{=\,\,}\beta_7 union_{it} + \beta_8 rur_{it} + \sum_t \beta_{9,t}year_{it} + a_i + u_{it} 
\end{align*}

### Pooled OLS
The pooled OLS model can be run from the `plm()` function as follows.
```{r}
est.pols <- plm(lwage ~ educ + black + hisp + exper + I(exper^2) + married + union + rur + year,
                data = df, index = c("id","year"), model = "pooling")
```

4. Interpret the coefficient on $\beta_7$ in the pooled OLS model

### Random effects
```{r}
est.re <- plm(lwage ~ educ + black + hisp + exper + I(exper^2) + married + union + rur + year, 
              data = df, index = c("id","year"), model = "random")
```

5. Explain why the standard errors in the random effects model are larger (or equal to) those of the pooled OLS model.

6. What is the estimate of $\theta$ in the RE model? (Hint: check `est.re$ercomp$theta`) What does this tell you about what you expect the random effects estimates to be relative to the fixed effects estimates?

### Fixed effects
```{r}
est.fe <- plm(lwage ~ I(exper^2) + married + union + rur + year, 
              data = df, index = c("id","year"), model = "within")
```

7. Explain why we cannot estimate coefficients on `educ`, `black`, `hisp`, or `exper` in the fixed effects model. (Note: the reason for not being able to estimate `exper` is more nuanced)

## Clustered standard errors
As discussed in class, the most appropriate standard errors are not those reported by `plm()`. Let's compute standard errors that account for within-person serial correlation, and that are robust to heteroskedasticity:
```{r}
clust.po <- coef_test(est.pols, vcov = "CR1", cluster = "individual")
clust.re <- coef_test(est.re,   vcov = "CR1", cluster = "individual")
clust.fe <- coef_test(est.fe,   vcov = "CR1", cluster = "individual")
```

We can put these all in one `stargazer` table:
```{r}
stargazer(est.pols,est.re,est.fe,se=list(clust.po$SE,clust.re$SE,clust.fe$SE),type="text")
```

8. Interpret the coefficient on `union` across the three models.

9. Comment on the comparability of the pooled OLS and RE estimators now that within-person serial correlation has been addressed.
