---
title: "In-Class Lab 14"
author: "ECON 4223 (Prof. Tyler Ransom, U of Oklahoma)"
date: "March 19, 2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', fig.keep = 'none')
```

The purpose of this in-class lab is to use R to practice with two-stage least squares (2SLS) estimation. The lab should be completed in your group. To get credit, upload your .R script to the appropriate place on Canvas.

## For starters
Open up a new R script (named `ICL14_XYZ.R`, where `XYZ` are your initials) and add the usual "preamble" to the top:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Add names of group members HERE
library(tidyverse)
library(wooldridge)
library(broom)
library(AER)
library(magrittr)
library(stargazer)
```

### Load the data
We're going to use data on working women.

```{r}
df <- as_tibble(mroz)
```

### Summary statistics
Like last time, let's use `stargazer` to get a quick view of our data:
```{r}
df %>% as.data.frame %>% stargazer(type="text")
```

1. Is it a problem that `wage` and `lwage` have 428 observations, but all of the other variables have 753 observations?

### Drop missing wages
Using the `filter()` and `is.na()` functions, drop the observations with missing wages. (I suppress the code, since you should know how to do this.)

```{r include=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
df %<>% filter(!is.na(wage))
```

## The model
We want to estimate the return to education for women who are working, using mother's and father's education as instruments:
\[
\log(wage) = \beta_0 + \beta_1 educ + \beta_2 exper + \beta_3 exper^2 + u
\]
where $wage$ is the hourly rate of pay, $educ$ is years of education, and $exper$ is labor market experience (in years).

### First stage regression
Let's estimate the first stage regression, which is a regression of the endogenous variable ($educ$) on the instrument(s) ($motheduc$ and $fatheduc$) and the exogenous explanatory variables ($exper$ and $exper^2$).[^1]

Run this regression (again, I suppress the code). Call the estimation object `est.stage1`.

```{r include=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
est.stage1 <- lm(educ ~ motheduc + fatheduc + exper + I(exper^2), data=df)
```

2. Double check that $motheduc$ and $fatheduc$ are jointly significant with an F-stat larger than 10:
```{r}
linearHypothesis(est.stage1,c("motheduc","fatheduc"))
```

### Second stage regression
In the second stage, we estimate the log wage equation above, but this time we include $\widehat{educ}$ on the right hand side instead of $educ$, where $\widehat{educ}$ are the fitted values from the first stage.

In R, we can easily access the fitted values by typing `fitted(est.stage1)`.

Let's estimate the second stage regression:
```{r}
est.stage2 <- lm(log(wage) ~ fitted(est.stage1) + exper + I(exper^2), data=df)
```

### Both stages at once
The standard errors from the above second stage regression will be incorrect.[^2] Instead, we should estimate these at the same time. We can do this with the `ivreg()` function, just like in the previous lab.
```{r}
est.2sls <- ivreg(log(wage) ~ educ + exper + I(exper^2) | motheduc + fatheduc + exper + I(exper^2), data=df)
```

3. Estimate the OLS model (where $educ$ is not instrumented). Then compare the output for all three models (OLS, 2SLS "by hand", 2SLS "automatic").
```{r include=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
est.ols <- lm(log(wage) ~ educ + exper + I(exper^2), data=df)
```
```{r}
stargazer(est.ols,est.stage2,est.2sls, type="text")
```

4. Comment on the IV estimates. Do they make sense, relative to what we think would bias the returns to education? Is the exogeneity condition on $motheduc$ and $fatheduc$ plausible?

[^1]: Note that you can easily include the quadratic in experience as `I(exper^2)` without having to create this variable in a `mutate()` statement.
[^2]: The reason is that error term in the second stage regression includes the residuals from the first stage, but the standard errors fail to take this into account.
