---
title: "In-Class Lab 8"
author: "ECON 4223 (Prof. Tyler Ransom, U of Oklahoma)"
date: "February 12, 2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', fig.keep = 'none')
```

The purpose of this in-class lab is to practice conducting ***joint*** hypothesis tests of regression parameters in R. We will do this using t-tests and F-tests. The lab should be completed in your group. To get credit, upload your .R script to the appropriate place on Canvas. 

## For starters
Open up a new R script (named `ICL8_XYZ.R`, where `XYZ` are your initials) and add the usual "preamble" to the top:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Add names of group members HERE
library(tidyverse)
library(broom)
library(wooldridge)
library(car)
library(magrittr)
```

### Load the data
We'll use a data set on earnings and ability, called `htv`. The data set contains a sample of 1,230 workers.
```{r}
df <- as_tibble(htv)
```

Check out what's in the data by typing
```{r}
glimpse(df)
```

The main variables we're interested in are: wages, education, ability, parental education, and region of residence (`ne`, `nc`, `west`, and `south`).

### Create regional factor variable
Let's start by creating a factor variable from the four regional dummies. Borrowing code from lab 6, we have:
```{r}
df %<>% mutate(region = case_when(ne==1 ~ "Northeast",
                                  nc==1 ~ "NorthCentral",
                                  west==1 ~ "West",
                                  south==1 ~ "South")) %>%
        mutate(region = factor(region))
```

## Regression and Hypothesis Testing
Estimate the following regression model:
\[
educ = \beta_0 + \beta_1 motheduc + \beta_2 fatheduc + \beta_3 abil + \beta_4 abil^2 + \beta_5 region + u
\]
Note that $abil$ is in standard deviation units. You will need to use a `mutate()` function to create $abil^2$ (not shown here). Call it `abilsq`. $region$ represents the factor variable you created above.[^1]
```{r include=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
df %<>% mutate(abilsq = abil^2)
```
```{r}
est <- lm(educ ~ motheduc + fatheduc + abil + abilsq + region, data=df)
tidy(est)
```
### t-test
1. Test the hypothesis that $abil$ has a linear effect on $educ$.

### F-test (single parameter)
2. Now test that $motheduc$ and $fatheduc$ have equal effects on $educ$. In other words, test $H_0: \beta_1=\beta_2; H_a: \beta_1 \neq \beta_2$. To do this, you will need to obtain $se(\beta_1 - \beta_2)$. Luckily, R will do this for you with the `linearHypothesis()` function in the `car` package:
```{r}
linearHypothesis(est, "motheduc = fatheduc")
```
The resulting p-value is that of an F test, but one would get an identical result by using a t-test, since this is a simple hypothesis (see @wooldridge, pp. 125-126).

### F-test (multiple parameters)
The p-values from the previous regression might indicate that the three region dummies don't contribute to education. 

3. Test the hypothesis that they don't; i.e. test 
\[
H_0: \text{all region dummies}=0; \\
H_a: \text{any region dummy}\neq 0
\]

The code to do this again comes from the `linearHypothesis()` function. The syntax is to encolose each component hypothesis in quotes and then surround them with `c()`, which is how R creates vectors.
```{r}
linearHypothesis(est, c("regionNortheast=0", "regionSouth=0", "regionWest =0"))
```
or, more simply,
```{r}
linearHypothesis(est, matchCoefs(est,"region"))
```

Alternatively, you can perform the F-test as follows (no need to put this in your R-script; I'm just showing you how to do it "by hand"):
```{r}
est.restrict <- lm(educ ~ motheduc + fatheduc + abil + abilsq, data=df)
Fstat.numerator   <- (deviance(est.restrict)-deviance(est))/3
Fstat.denominator <- deviance(est)/1222
Fstat <- Fstat.numerator/Fstat.denominator
p.value <- 1-pf(Fstat,3,1222)
```
This gives the exact same answer as the `linearHypothesis()` code.

# References

[^1]: Here the notation of $\beta_5 region$ is not quite right. It more technically should be written $\beta_5 region.NE + \beta_6 region.S + \beta_7 region.W$, where each of the $region.X$ variables is a dummy. The way it is written above, $\beta_5 region$ implies that $\beta_5$ is a vector, not a scalar.
